---
topic: "sustainability"
title: "Environmental Product Responsibiliy indicator"
output: envreportutils.internal::print_ver
params:
  draft: TRUE

---

# Background 

The dataset provide as part of environment reporting legislation requirements in pdf format and has been transposed to an excel format. Detailed information about the program can be found [here](https://www2.gov.bc.ca/gov/content/environment/waste-management/recycling/product-stewardship).

This indicator will be replacing the [tire recycling indicator](https://www2.gov.bc.ca/gov/content/environment/research-monitoring-reporting/reporting/environmental-reporting-bc/sustainability-indicators).



```{r echo = FALSE, include=params$draft, results='asis'}
cat("\\begin{center}
\\fbox{
  \\begin{minipage}[c]{14cm}
 
   \\begingroup\\centering{\\textbf{DRAFT – NOT FOR DISTRIBUTION}}
   
    \\endgroup
    
    \\vspace{1mm}
   
    \\text{This DRAFT public reporting product is under review and may change. It is not intended}
    \\text{to be shared, cited or used at this time. Please contact Environmental Reporting BC program}
    \\text{staff if you have questions about the content. }
   
  \\end{minipage}
}
\\end{center}")
```






```{r echo=FALSE, message=FALSE}
library(ggplot2)
ggplot(mtcars, aes(x = hp, y = mpg)) + 
  geom_point() + 
  geom_smooth()
```

Cum vivamus penatibus sed mus fringilla non sociosqu taciti ut. Sed, semper lacus justo ultricies penatibus praesent. Tortor vivamus fermentum libero, libero urna montes orci luctus ligula? Tellus et mauris sed a in lacinia imperdiet. Sed nec penatibus sed at mauris etiam. Interdum morbi arcu vitae lectus cubilia eros, imperdiet felis. Nisl, lacinia vel, nisl, rutrum ac, elementum felis. Dui ante eros ex mi, interdum velit accumsan sociosqu, tempus. Erat vulputate potenti nascetur. Malesuada, curabitur, leo himenaeos, molestie euismod sit. In vitae sollicitudin adipiscing commodo.

## References and Other Useful Links

* Link 1
* Link 2

## Data

\*By accessing these datasets, you agree to the licence associated with each file, as indicated in parentheses below.

* Data 1
* Data 2

----

Published and Available On-Line at Environmental Reporting BC (Month Year):  
[](http://www.env.gov.bc.ca/soe)

Email correspondence to: envreportbc@gov.bc.ca

Suggested Citation:  
Environmental Reporting BC. Year. Indicator Title. State of Environment Reporting, Ministry of Environment and Climate Change Strategy, British Columbia, Canada.
