---
title: "Project_Notes"
author: "G Perkins"
date: "June 5, 2019"
output: 
  html_document: 
    toc : true
    toc_float: 
      collapsed: false
      smooth_scroll: false
      
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# EPR: Project Notes

The dataset provide as part of environment reporting legislation in pdf format and has been transposed to an excel. Detailed informatio about the program can be found here <https://www2.gov.bc.ca/gov/content/environment/waste-management/recycling/product-stewardship> 

This indicator will be replacing the tire recycling indicator<https://www2.gov.bc.ca/gov/content/environment/research-monitoring-reporting/reporting/environmental-reporting-bc/sustainability-indicators>


## Initial data summary: 

Main indicators: 

- Beverage containers 
- Oil 
- Tires
- Paint-Flam_Pest
- Elect
- Lead-Acid Batteries 
- Packaging and Printed Paper


Along with the BC population measures as of populations for 2018 and program financials and population estimates. 

# Beverage containers. 
Collected from multiple groups: Encorp Pacific (Canada), BC Brewers Recycles Collection Council. Both of these collect data on the units collected, weight collection (tonnes) with population data from 2006 - 2017. There is also some financial data ( revenue/ expenditure /deposits charged / units moved) and other (CO2 Equivalent Reduced (tonnes), weight of material)

Data exploration points to consider:  

- group Encorp and BCBRCC data together by Abcolute collection units Collected per regional district 
- BCBRC years may be different. 
- use popultaion rates reported in BCBRCC as first pass. (seperate csv)
- timing diffrence due to 2007/08	2008/09	2009/10 BCBRCC reported as financial years to March 31st 
- some merged cells may be problemtic in xlsx/csv processing 
- comox/strathcona split into two regions in 2011 onwards only in the Encorp data set - recorded seperately for BRCC.
- liquor distribution branch ( 2000 - 2006) only active for short time. 
- Error in 2017 data set in per capita returns for BRCC, better to work off the raw data sets. 

Some notes of caution

- ** FINANCE NOTE: Not all programs report expenditure, and information is not available for all years. These figures represent 'reported' expenditure only (i.e. that included in published annual reports) and should not be taken as total costs of EPR programs.	
- ** OTHER CATERGORY NOTE: Reduced Pollutants refers to pollutants that would have been produced if BLD's recycling system did not exist, and the bottles they would have recycled were instead replaced by containers made from virgin materials.	


Some general conclusions 

- Data can be split in geographic (by regions) or a-spacial (BC wide). 
- Regional data: units and tonnes of recycled material & combine with population
- BC wide data: Financial metrics/ units moved / other (tonnes of emisions etc. )


- Key information to work with raw values, finance data and units, 
- Finance data is BC wide, 
- recycling data is regional based. 



## Manual processsing
convert to csv, add company name column and a measure catergory (priority measure, financial, Units Moved,other )
create a seperate population cvs to be used with the other recycling metrics. 

- create a single pop csv to use in the analysis, error in raw data (population for Caribou missing 10,000)

## Data exploration 
read in the data

```{r loadlibs, echo=FALSE,include = FALSE}

x <- c("dplyr","ggplot2","tidyr","stringr","reshape", "bcmaps", "sf") 
lapply(x, library, character.only = TRUE) ; rm(x)  # load the required packages

```

```{r loaddata, echo=TRUE,include = TRUE}
## Load  data files
setwd("C:/Temp/Github/recycling-indicator/")
data.dir <- "data/"
rdata<- read.csv(paste(data.dir,"bev.csv",sep = ""))
source('00_Functions.R')

#head(rdata)

# extract the population data and export as csv
pop <- rdata %>% dplyr::filter(Measure == 'Population-') %>%
         dplyr::select(-c(Company,Catergory,Measure))

pop.long <- pop %>% gather("year", "n",2:19)
pop.long$year = sub("X","",pop.long$year)  # get rid of x on year column
pop.long$pop <- replaceCommas(pop.long$n) # fix the fromatting in numeric
head(pop.long)


```

## Explore the non spatial data (all regions combined): Financial, units mooved and other
1) Financial 

```{r financial, echo=TRUE,include = TRUE}
# Grab the financial data
fdata <- rdata %>% dplyr::filter(Catergory == 'Financial') %>%
          dplyr::select(-c(Catergory,Region)) %>%
          dplyr::filter(!Company == 'EncorpPacific_BRCCC') %>%
          gather("year", "n",3:20)
#fdata$n = sub("$","",fdata$n)  # get rid of x on year column
fdata$n <- replaceDollars(fdata$n) # fix the fromatting in numeric
fdata$n <- replaceCommas(fdata$n) # fix the fromatting in numeric
fdata$n.m <-fdata$n/1000000
fdata$year = sub("X","",fdata$year)  # get rid of x on year column

sum.fdata <- fdata %>% group_by(Measure,year) %>%
  summarise(total = sum(n.m,na.rm = TRUE))
```

### How does the amount of revenue influence the number of unclaimed deposits?

```{r unclaimed, echo=TRUE,include = TRUE}
to.keep <- c('Unclaimed Deposits','Expenditure-Consumer Awareness')

sum.fdata <- sum.fdata %>%
  filter(Measure %in% to.keep )
# Does spending more on consumer awareness decrease unclaimed deposits
    ggplot(sum.fdata,aes(year,total,fill=Measure)) +
      geom_bar(stat="identity",position="dodge") +
      labs(title="Unclaimed deposits and consumer-expenditure", x = "Year", y = " Amount ($1,000,000)")
    #ggsave(paste('out/',"01_Beverage_UnitsMoved.png"))

```

There is not a strong correlation between with awareness, may need to adjust this per unit? or check if the awareness expenditure is for multiple programs (ECPOR and BCRCCC)

2) Units Moved
Is there a difference in the units moved (recoved or sold over time?)

```{r Units moved , echo=TRUE,include = TRUE}
udata <- rdata %>% dplyr::filter(Catergory == 'Units Moved') %>%
  dplyr::select(-c(Catergory,Region))%>%
  dplyr::filter(!Company == 'EncorpPacific_BRCCC') %>%
  dplyr::filter(!Measure == 'Recovery Rate (%)  Regulation Target 75%') %>%
  gather("year", "n",3:20)

udata$n <- replaceCommas(udata$n) # fix the fromatting in numeric
udata$n.m <-udata$n/1000000
udata$year = sub("X","",udata$year)  # get rid of x on year column

# summarise per year
sum.udata <- udata %>% group_by(Measure,year) %>%
  summarise(total = sum(n.m,na.rm = TRUE))

  # make a pretty graph
  ggplot(sum.udata,aes(year,total,fill=Measure)) +
    geom_bar(stat="identity",position="dodge") +
    labs(title="Recycled Units Returned and Sold", x = "Year", y = "Total no. (millions)")
   # ggsave(paste('out/',"01_Beverage_UnitsMoved.png"))

 
```
   
   Appears to be a gap in the 2007 and 2008 year that needs to be checked as units show large decrease. 
   Perhaps better to show the recovery rate, rather than recoved and sold (two verus one number)
   
```{r Recovery rate  , echo=TRUE,include = TRUE}    
    #calculate the recovery rate and make a line plot
sum.udata1 <-  udata %>% group_by(Measure,year) %>%
  summarise(total = sum(n.m,na.rm = TRUE)) %>%
  spread(., Measure,total)

sum.udata1$RecoveryRate = sum.udata1$'Units Recovered' / sum.udata1$'Units Sold' *100
sum.udata1 <- sum.udata1[-1,]

    ggplot(sum.udata1, aes(x = year, y = RecoveryRate))+
      geom_point() +
      ylim(60,100) +
      geom_hline(yintercept = 75, color = "red", lty = 2) +
      labs(title="Recycled Units Recovery Rate (%)", x = "Year", y = "Recovery Rate %")
      #ggsave(paste('out/',"02_Beverage_UnitsMoved.png"))

```

3. Other catergory. 
This includes data on the amount of emmission "saved/reduced" and total volumes 

```{r, echo = TRUE}
odata <- rdata %>% dplyr::filter(Catergory == 'Other') %>%
    dplyr::select(-c(Company, Catergory,Region)) %>%
    gather("year", "n",2:19)
odata$n <- replaceCommas(odata$n) # fix the fromatting in numeric
odata$n.t <-odata$n/1000
odata$year = sub("X","",odata$year)  # get rid of x on year column
odata$Measure = sub("tonnes","Tonnes",odata$Measure)  # get rid of x on year column

odata <- odata %>% group_by(Measure,year) %>%
  summarise(total = sum(n.t,na.rm = TRUE))

ggplot(odata, aes(year,total,fill=Measure))+
  geom_bar(stat="identity",position="dodge") +
  labs(title="Tonnes of material and reduction of pollutants", x = "Year", y = "Tonnes (thousands)") +
  scale_y_continuous(limits = c(0,250))

```
 
This plot could be reduced or split into two plots emissions saved and emissions and potentially weight of material recycled? 

## Explore the spatial data sets: Units and weight per capita
# 
This includes regional based estimate. Still need to edit the comox strathcoma regions. Also some data entry errors are apparent when calculating the general plots 

```{r explor spatial dataset, echo = TRUE , warn=1}
# extract the raw unit data and add popultaion data 
pdata <- rdata %>% dplyr::filter(Catergory == 'Priority Measures') %>%
  dplyr::select(-c(Catergory)) %>%
  dplyr::filter(!Company == 'EncorpPacific_BRCCC')%>%
  dplyr::filter(Measure %in% c("Absolute Collection-Units Collected-","Absolute Collection-Weight Collected (Tonnes)-")) %>%
  gather("year", "n",4:21)
# format the data tyep 
pdata$year = sub("X","",pdata$year)  # get rid of x on year column
pdata$n <- replaceCommas(pdata$n) # fix the fromatting in numeric

# group together the multiple companies
sum.pdata <- pdata %>% group_by(Measure, Region, year) %>%
  summarise(total = sum(n,na.rm = TRUE))

# merge in the pop.long form data
ppdata <- left_join(sum.pdata,pop.long, by = c("Region","year")) %>%
  dplyr::select(-c('n')) %>%
  mutate(unit.per.cap = total / pop)
```
No w data is format split this into different units, tonnes and units per capita and create some summary plots 

```{r split into diffrent unirs and make plots, echo = T, results = T}
# Split data set into weight and units
units.per.cap <- ppdata %>% filter(Measure == 'Absolute Collection-Units Collected-' )
weight.per.cap <-  ppdata %>% filter(Measure == 'Absolute Collection-Weight Collected (Tonnes)-' )

    ## Units per capita
    p1 <- ggplot(units.per.cap,aes(year,unit.per.cap)) + facet_wrap(~Region)+
      geom_bar(stat="identity",position="dodge") +
      labs(title="Regional Units Recycled per capita", x = "Year", y = " units per capita")
   # ggsave(paste('out/',"04_Beverage_UnitsPerCap.png"))
    p1
```
This plots shows some obvious data errors to check. Also highlights need to adjust comox and stracthcona datasets as not being correctly calculated. 

```{r weights per capita, echo = T, results = T}
    ## weight per capita
    p2 <- ggplot(weight.per.cap,aes(year,unit.per.cap)) + facet_wrap(~Region)+
      geom_bar(stat="identity",position="dodge") +
      labs(title="Regional weight of recycling (tonnes) per capita", x = "Year", y = "weight per cap (tonnes")
    #ggsave(paste('out/',"04_Beverage_weightPerCap.png"))
    p2

```

Plot 2 also shows simialr issue with naming to be fixed 

We can also investigate the difference from the provincial average: 

```{r calculate average and }

# calculate the provincial average
bc.units.per.cap <- units.per.cap %>% na.omit() %>%
    group_by(year) %>%
    summarise(BCave = mean(unit.per.cap))

regional.units.per.cap <- units.per.cap %>% na.omit() %>%
    group_by(year,Region) %>%
    summarise(ave = mean(unit.per.cap))
# join the regional and prov. ave data and calculate the difference

diff.df <-left_join(regional.units.per.cap,bc.units.per.cap, by = 'year')
diff.df <- diff.df %>% mutate(delta = ave-BCave)
diff.df$response <- ifelse( diff.df$delta < 0, "below", "above")
    # calculate the deviation from average

    # Diverging Barcharts
    p4 <- ggplot(diff.df, aes(x=Region, y=delta,label=delta)) + facet_wrap(~year) +
      geom_bar(stat='identity', aes(fill=response), width=.5)  +
     scale_fill_manual(name="Mileage",
                        labels = c("Above Average", "Below Average"),
                        values = c("above"="#00ba38", "below"="#f8766d")) +
      labs(title= "Regional difference from the average BC units per capita recycling") +
      coord_flip()
    #ggsave(paste('out/',"05_regional_bc_ave_unitsPerCap_per_yr.png"))
    p4
    # Diverging Barcharts (all years)
    p5 <-  ggplot(diff.df, aes(x=Region, y=delta, label = delta)) +
      geom_bar(stat='identity', aes(fill=response), width=.5)  +
      scale_fill_manual(name="Mileage",
                         labels = c("Above Average", "Below Average"),
                          values = c("above"="#00ba38", "below"="#f8766d")) +
       labs(title= "Regional difference from the average BC units per capita recycling") +
      coord_flip()
    # ggsave(paste('out/',"06_regional_bc_ave_unitsPerCap_allyrs.png"))
    p5
```


Thoughts on Beverage recycling:

- Map regional data sets for presentation (leaflet)
- difference maps better than raw values ? Difference to average works with change per year indicated 
- could present some metrics within graphs (ie non - spatial data sets) 
- Northern Rockies missing population data - need to pull data from stats Canada per year 


# Lubraicating Oil, Filters and containers. 

Collection broken in regional and temporal data sets from 2010 to 2017 including: 

- Used oil (litres per person) 
- Used filters (kg per person) 
- Containers (kg per person)
- Antifreeze (kg person)

Data exploration points to consider: 
Temporal and a-spatial data sets include financial, collection sites, units moved, and end fate for products. 
Collection sites does not look like much change in trends over time?
Could be oportunity to combine a number of measure (ie beverage together). Comox and Strathcoma are treated seperatiely in this analysis. 

Could also combine the % recover rate over time for different metrics in a single graph 

```{r loaddata oils, echo=TRUE,include = TRUE}
## Load  data files
setwd("C:/Temp/Github/recycling-indicator/")
data.dir <- "data/"
rdata<- read.csv(paste(data.dir,"LubOilFilt.csv",sep = ""))

#head(rdata)

# extract the population data and export as csv
#pop <- rdata %>% dplyr::filter(Measure == 'Population-') %>%
#         dplyr::select(-c(Company,Catergory,Measure))

#pop.long <- pop %>% gather("year", "n",2:19)
#pop.long$year = sub("X","",pop.long$year)  # get rid of x on year column
#pop.long$pop <- replaceCommas(pop.long$n) # fix the fromatting in numeric
#head(pop.long)


```

## Explore the non spatial data (all regions combined): Financial, units mooved and other
1) Financial 

```{r financial oil, echo=TRUE,include = TRUE}
# Grab the financial data
#fdata <- rdata %>% dplyr::filter(Catergory == 'Financial') %>%
#          dplyr::select(-c(Catergory,Region)) %>%
#          dplyr::filter(!Company == 'EncorpPacific_BRCCC') %>%
#          gather("year", "n",3:20)
#fdata$n = sub("$","",fdata$n)  # get rid of x on year column
#fdata$n <- replaceDollars(fdata$n) # fix the fromatting in numeric
#fdata$n <- replaceCommas(fdata$n) # fix the fromatting in numeric
#fdata$n.m <-fdata$n/1000000
#fdata$year = sub("X","",fdata$year)  # get rid of x on year column

#sum.fdata <- fdata %>% group_by(Measure,year) %>%
#  summarise(total = sum(n.m,na.rm = TRUE))
```


